---
title: "Untitled"
author: "Matt Twigg"
date: "June 10, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction#
This practicum intends to look at business growth and predicting that growth

#Exploration#


```{r libraries}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(forcats)

```
load business
```{r 1}
business <- read.csv("Business_Entities_in_Colorado.csv", header = TRUE)
```
look at the head of the data
```{r 2}
head(business)
```
how many rows do I have
```{r 3}
#totBusiness <- nrow(business)
#totBusiness 
```
Let's reduce the columns
```{r 4}
b1 <- business %>% select(
  entityid, 
  entityname, 
  principaladdress1, 
  principalcity, 
  principalstate,
  principalzipcode,
  principalcountry, 
  entitystatus, 
  jurisdictonofformation, 
  entityformdate
  )
```
Let's see just the years of registration
```{r 5}
date <-  as.Date(b1$entityformdate,'%m/%d/%Y')
year <- as.numeric(format(date,"%Y"))
#View(year)
```
Let's add a year column to the reduced data set
```{r}
b2 <- cbind(b1, year)
b2
```
Finally, let's see basic growth of registered companies in Colorado for the past 135 years
```{r }
hist(b2$year, breaks = 135)
```

```{r}
b2$principalzipcode <- trimws(b2$principalzipcode, which = c("left"))
```

```{r}
b2$principalzipcode <- substr(b2$principalzipcode, start = 1, stop = 5)
```

```{r}
b2
```

```{r}
b2$principalzipcode <- as.numeric(b2$principalzipcode)
```


```{r}
hist(b2$principalzipcode)
```



What are the home states for registered businesses
```{r}
FreqCO <- unique(b2$principalstate)
FreqCO

```
Homestates that have registered businesses in Colorado and how many businesses
```{r}
test1 <- summary(b2$principalstate)
test1
```
Same data but in a table
```{r}
homeStates = table(b2$principalstate)
homeStates
```


```{r}
other <- totBusiness - 1672148
portions <- c(other, 1672148)
states <-c("other", "Colorado")
pie(portions,labels = states, main="Registered Businesses Home State")
```

```{r}
topHomeStates <-summary(fct_infreq(b2$principalstate))
topHomeStates
```

```{r}
class(homeStates)
```

```{r}
tester <- barplot(topHomeStates, main="Registered Business Home State", xlab = "States")
tester
```

```{r}

b3 <- b2 %>% separate(entityname, c("company", "dissolved"), sep = "Dissolved")
```

```{r}
b3
```

```{r}
b3$dissolved <- trimws(b3$dissolved, which = c("left"))
```


```{r}
testb3 <- b3 %>% separate(dissolved, c("month", "day", "yrdslv"), sep = " ")
datedslv <-  as.Date(testb3$yrdslv,'%Y')
yeardslv <- as.numeric(format(datedslv,"%Y"))
```

```{r}
b4 <- cbind(b3, yeardslv)
b4
```

```{r}
hist(b4$yeardslv, breaks = 135)
```


```{r}
sum(is.na(testb3$yrdslv))
```


```{r}
linearMod <- lm(dist ~ speed, data=b3)
```


```{r}
#disornot <- grepl('Dissolved', b3)
#View(disornot)
```

```{r}
b2 <- cbind(b2, disornot)
head(b2)
```



```{r}
b4 %>%
  separate(entityname, c("companyname", "disyear"), sep="\\s+(?=\\S*$)")
```
```{r}
b5 <- b4 %>%
  separate(entityname, c("companyname", "disyear"), sep="\\s+(?=\\S*$)")
```
```{r}
b5
```

```{r}
year
yeardissolved <- as.numeric(gsub("[^0-9.-]", "", b5$disyear))
```

```{r}
yeardissolved <- trimws(yeardissolved, which = c("both"))
yeardissolved
```

```{r}
typeof(yeardissolved)
```

```{r}
sapply(b4, class)
```

```{r}
b6 <- cbind(b4, yeardissolved)
head(b6)
```

```{r}
b6$yeardissolved[is.na(b6$yeardissolved)] <- 1880
head(b6)
```

```{r}
str(b6)
```




find population in zip code
home price
multi linear
stock market data snp 500

